---
title: "Fifa Football Analysis"
author: "DEON S"
date: "14/08/2022"
output: html_document
---

#### Reading the csv file

```{r}
fifa <- read.csv("C:/Users/LENOVO/Dropbox/PC/Desktop/Project 2/fifa_cleaned.csv")
```

#### To view the imported csv file

```{r}
View(fifa)
```

### Into Data Cleaning
#### To know the size of data

```{r}
dim(fifa)
```

#### summary statistics for all the columns of the data
```{r}
summary(fifa)
str(fifa)
```

#### To check null values
```{r}
any(is.na(fifa))
```

#### To check the count of null values
```{r}
sum(is.na(fifa))
```

#### To get the specific column with null values
```{r}
sum(is.na(fifa$TotalCharges))
```
```{r}
colSums(is.na(fifa))
```

#### To remove columns 
```{r}
fifa_1 <- fifa[-c(1,20,24,25,27:30,60:92)]
```

#### To remove the null values
```{r}
fifa_edit <- na.omit(fifa_1)
```

#### is there still a null value
```{r}
sum(is.na(fifa_edit))
```

#### To check null values
```{r}
any(is.na(fifa_edit))
```

#### Further more data pre-processing
```{r}
fifa_1 <- fifa[-c(23,25,26)]
```

#### to export clean dataset
```{r}
write.csv(fifa_edit, "C:/Users/LENOVO/Dropbox/PC/Desktop/Project 2/fifa_clean.csv")
```

---------------------------------------------------------------------------------------------------
#### All the irrelevant data was cleaned and copied to a new csv file named "fifa_final" and further data analysis is done based on that.
```{r}
fifa_final<-read.csv("C:/Users/LENOVO/Dropbox/PC/Desktop/Project 2/fifa_final.csv")
View(fifa_final)
```

```{r}
dim(fifa_final)
```

#### To check the datatypes
```{r}
str(fifa_final)
```
#### loading necessory packages

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(heatmaply)
library(ggcorrplot)
library(devtools)
library(easyGgplot2)
library(caTools)
library(ROCR) 
library(leaps)
```

#summary statistics for all the columns of the data
```{r}
summary(fifa_final)
str(fifa_final)
```
#To check for null values
```{r}
which(is.na(fifa_final))
```

#### numeric datatypes
```{r}
fifa_numeric<-fifa_edit[c('age', 'height_cm', 'weight_kgs', 'overall_rating', 'potential', 'value_euro', 'wage_euro', 'international_reputation.1.5.', 'weak_foot.1.5.', 'skill_moves.1.5.', 'club_rating', 'crossing', 'finishing', 'heading_accuracy', 'short_passing', 'volleys', 'dribbling', 'curve', 'freekick_accuracy', 'long_passing', 'ball_control', 'acceleration', 'sprint_speed', 'agility', 'reactions', 'balance', 'shot_power', 'jumping', 'stamina', 'strength', 'long_shots', 'aggression', 'interceptions', 'positioning', 'vision', 'penalties', 'composure', 'marking', 'standing_tackle', 'sliding_tackle')]

```

#### categorical datatypes
```{r}
fifa_categorical<-fifa_edit[c('name', 'full_name', 'birth_date', 'positions', 'nationality', 'preferred_foot', 'work_rate', 'body_type', 'club_team', 'club_position', 'contract_end_year')]

```

#### EDA ***********************************************************

#### count of height of the players
```{r}
ggplot(fifa_final, aes(x = height_cm)) +
  geom_bar(
    
  )

```

#### Heatmap of correlation between the numeric variables
```{r}
ggcorrplot::ggcorrplot(cor(fifa_numeric))
```

#### To check the unique body types
```{r}
unique(fifa_final[c("body_type")])

```
#### scatterplot of wage in euro vs height of the player
```{r}
ggplot2.scatterplot(data=fifa_final, xName='wage_euro',yName='height_cm', 
                    groupName='body_type', mapping=aes(size = body_type),
                    backgroundColor="white", setColorByGroupName=TRUE) 
```

#### scatterplot of value in euro vs height of the player
```{r}
ggplot2.scatterplot(data=fifa_final, xName='value_euro',yName='height_cm', 
                    groupName='body_type', mapping=aes(size = body_type),
                    backgroundColor="white", setColorByGroupName=TRUE) 
```

#### chi-square test to know whether two categorical variables are dependent or not
```{r}
chisq.test(fifa_final$positions,fifa_final$preferred_foot)
```
```{r}
value_df <- fifa_final %>% dplyr::select(-name, -nationality, -club_team, -body_type, 
                                            -club_position)

value_df
```

#### player rating vs value in euro
```{r}
fifa_final %>% ggplot(aes(x=overall_rating,y=value_euro)) + geom_point() + geom_jitter() +
  labs(x = "player rating", y = "player value (euro)", 
       title = "Plot of player rating vs player value")
```

#### Histogram of height of the players
```{r}
hist(fifa_final$height_cm,xlab='Height',ylab='Frequency',main='Height of the players')
```

#### To check the unique positions
```{r}
df_position <- unique(fifa_final[c("positions")])
```

```{r}
View(df_position)
```

#### display the counts of df_position
```{r}
n_distinct(fifa_final$positions)
```
##### contains 204 unique postions

#### Make the column "Positions" have only one data per row
```{r}
fifa_final$positions <- gsub(" ", "", substr(fifa_final$positions,1,3))
fifa_final$positions <- gsub(",", "", substr(fifa_final$positions,1,3))
```

#### to check the unique Positions again
```{r}
df_position <- unique(fifa_final[c("positions")])
```

```{r}
View(df_position)
```

```{r}
n_distinct(fifa_final$positions)
```

#### Group the positions in GK, DEF, MID,FWD
```{r}
x <- as.factor(fifa_final$positions)
levels(x) <- list(GK =c("GK"),
                  DEF =c("LWB","LB","CB","RB","RWB"),
                  MID = c("LW","LM","CDM","CM","CAM","RM","RW"),
                  FWD = c("CF","ST"))
```

```{r}
df <-mutate(fifa_final,position = x)
```

#### viewing the data again
```{r}
View(x)
```

#### distribution of players based on the age
```{r}
g_age <- ggplot(data = df, aes(age))
g_age + 
  geom_histogram(col="orange", aes(fill = ..count..)) + ggtitle("Distribution based on Age")
```

#### relation between the Age of the players and their general playing position.
```{r}
g_age + 
  geom_density(col="orange", aes(fill = position), alpha=0.5) + facet_grid(.~position) + 
  ggtitle("Distribution based on Age and Position")
```
#### players overall rating
```{r}
g_overall <- ggplot(data = df, aes(overall_rating))
g_overall + 
  geom_histogram(col="orange", aes(fill = ..count..)) + ggtitle("Distribution based on Overall Rating")
```
#### Using the quantile function in R to obtain the top 1 % count of the player Value and Wage 
```{r}
top_1_percent_wage   <- quantile(df$wage_euro, probs=0.99)
filtered_wage <- filter(df, wage_euro > top_1_percent_wage)
```

```{r}
g_value <- ggplot(filtered_wage, aes(wage_euro))
g_value + 
  geom_histogram(aes(fill=..count..)) + 
  ggtitle("Distribution of top 1% value")
```

A large number of players are worth 25,000,000 with the count decreasing sharply as the price increases.

```{r}
top_1_percent_value   <- quantile(df$value_euro, probs=0.99)
filtered_value <- filter(df, value_euro > top_1_percent_value)
```

```{r}
g_wage <- ggplot(filtered_value, aes(value_euro))
g_wage + 
  geom_histogram(aes(fill=..count..)) + 
  ggtitle("Distribution of top 1% Value")
```
### Building the Multiple Regression model

#### # overall rating vs. value
```{r}
fit_overall <- lm(log(value_euro) ~ overall_rating, data = fifa_final)
summary(fit_overall)
```
 The model does not give a good accuracy. Therefore we proceed multiple regression with other set of variables.
 
#### potential vs. value
```{r}
fit_potential <- lm(log(value_euro) ~ potential, data = fifa_final)
summary(fit_potential)

fifa_final %>% ggplot(aes(x=potential,y=log(value_euro))) +
  geom_point() + geom_smooth(method='lm', formula= y~x) +
  labs(x="player potential", title = "Plot of potential vs. player value")
```

#### weekly wage_eur vs. value
```{r}
fifa_final %>% ggplot(aes(x=wage_euro,y=value_euro)) +
  geom_point() + geom_smooth(method='lm', formula= y~x)
```

#### multiple regression model by keeping value_euro as the dependent variable
```{r}
lm.fit.value = lm(value_euro ~
                    age+overall_rating+potential+wage_euro+
                    international_reputation.1.5.+
                    stamina+sliding_tackle+standing_tackle+height_cm, data = fifa_final)
summary(lm.fit.value)
```

Our multiple regression model gave us a accuracy of 0.79 which is quite good.


